### 2024-02-12 paper/paper.qmd
Results section completed and compiled with additions to Appendix too. usage.txt updated with one more prompt that was used. Discussion section left.

---
title: "Why Did Russia Fare So Poorly during COVID: An Analysis of Why One of the Best Prepared Countries Could Not Control Excess Deaths"
author: 
  - Aryaman Sharma
  - Aviral Bhardwaj
  - Janel Gilani
thanks: "Code and data are available at: https://github.com/Ary4m3n/covid-effect-russia.git; Replication on Social Science Reproduction platform available at: "
date: today
date-format: long
abstract: "Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aenean tempus libero sit amet nisl dapibus, varius dictum magna pellentesque. In lacinia nec ligula eget condimentum. Aenean imperdiet enim sit amet felis volutpat, eu ornare ante pretium. Donec ullamcorper tempus facilisis. Mauris porttitor libero ultrices tortor facilisis, quis varius sapien ultricies. Pellentesque et metus vel quam mollis dignissim ultricies a dui. Cras sodales porttitor pulvinar. Proin vel nunc sit amet sapien venenatis pellentesque nec at felis."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(ggplot2)
library(janitor)
library(dplyr)
library(rnaturalearth)
library(sf)
library(knitr)
```

# Introduction {#sec-introduction}

The COVID-19 pandemic has led to an enormous loss of life over the world and has presented an unprecedented challenge to public health, foreign policy and has additionally had an adverse effect on countries' economies around the world, over the past 5 years [@citeWHOeffect]. There have been over 774 million reported confirmed cases of the virus being contract and the outbreak has caused over 7 million deaths, which are claimed to be "under counted" and "under reported", suggesting that the actual number is even higher around 18.2 - 33.5 million based on the excess mortality estimates [@citeWikiCOVID]. This excessively high number of deaths has alerted the world to assess how countries assumed to be well-prepared to tackle such outbreaks were affected to this extent.

There were a subset of countries, namely India, the United States of America, Brazil, Russia and Mexico which were impacted the most in terms of having the highest cumulative total deaths [@citeThinkGlobalHealth]. According to the global data, Russia's COVID-19 related deaths since the start of the pandemic have been over 230,000. Russia also holds the highest mortality rate in Europe and the second-highest rate in Asia [@citeCarnegieRussia]. For these reasons and more, Russia and its residents have suffered a lot over the past few years, wherein the adverse impact of the outbreak has lead to a plethora of impacts on the country's economy and standard of living.

This paper will follow a reproduction of Jennifer B. Nuzzo and Jorge R. Ledesma's paper [@citeRepPaper] and findings and apply a Russian-facing lens to discuss the impact of the COVID-19 outbreak on Russia and what could have been done better and can be improved for the future in order for the impact of any upcoming pandemics can be lessened. Our paper aims to replicate their claim of how well-prepared countries did so poorly during the COVID-19 outbreak, where the Nuzzo and Ledesma have concentrated solely on the United States but in this paper we look at Russia with the same regards and analyse the impact of the pandemic on the death toll, and finally provide for ways to improve the laws, regulations and plans to tackle such outbreaks in the future. Our reproduction found out that how ever well prepared Russia seemed to be, it experienced an unprecedented rise in the death rate compared to other countries, and its excess death rate was one of the highest amongst all countries. This leads us to make an important judgement that better planning and regulations can help reduce the impact of such outbreaks in the future, which will be studied in this paper.

This paper is structured using the following sections: Data, Results and Discussion. In the Data (@sec-data) section, the data source of the datasets from the paper being replicated [@citeRepPaper] is discussed and the measurement and data cleaning process is outlined. In the Results (@sec-results) section, the paper summarizes the data findings and presents relevant maps and plots in order to study the trends. The paper ends with the Discussion (@sec-discussion) section, where the findings of the paper have been analysed and delved deeper into by studying the measures that can be taken to prevent the country from experiencing such an extensive effect of future pandemics. Additionally, the limitations and a further scope for the paper has been discussed here.

# Data {#sec-data}

As mentioned above, the data analysed in this paper is from the reproduction package of the paper by Nuzzo and Ledesma [@citeRepPaper]. The data was cleaned and analysed using the open source `R` programming language [@citeR]. R libraries and packages such as `tidyverse` [@citeTidyverse], `janitor` [@citeJanitor], `ggplot2` [@citeGgplot], `knitr` [@citeKnitr], `readr` [@citeReadr], `dplyr` [@citeDplyr], `rnaturalearth` [@citeRnaturalearth] and `sf` [@citeSf]. In the following sections, we will discuss the raw data in (@sec-data-raw), the data cleaning process in (@sec-data-clean) and then move on to discussing the measurement in (@sec-data-measurement).

## Raw Data {#sec-data-raw}

This paper will replicate the data that was obtained from the replication package of the paper [@citeRepPaper]. Specifically, in the paper we use 3 datasets that will help us analyse the claim of Russia not being able to control the wave of the COVID-19 pandemic well. In general, there is a lot of data available on the effect of COVID-19 on countries, however, all the data we needed for this paper was included in the paper by Nuzzo and Ledesma.

The first raw dataset includes data on the Global Health Security index for all countries. The GHS index is an indicator of how well prepared or "secure" a country is against epidemics or pandemics like the COVID-19 outbreak. This dataset has a plethora of data on indices that are not relevant to us for this paper. Additionally, the dataset contains data for years 2019 and 2021. For our analysis, we only require data for the GHS index of countries for 2021.

The second raw dataset includes data on all cause death rates for all countries from 1960-2021. This dataset has loads of information that was mainly unnecessary to us for this paper, mainly the data from 1960-2010 which is not required as we are only interested in looking at the effect of COVID-19 which is mainly data from 2020-2021. However, in order to study the trend, we will use data for all cause death rates for countries around the world from 2010-2021.

The third raw dataset is from the Institute of Health Metrics and Evaluation (IHME) which presents data on the COVID death rate, excess deaths, excess death rate and ratio of excess death rate over COVID death rate for all countries in the world. Again, this dataset too contained unnecessary information that was of no need to us which the cleaned data process outlined in @sec-data-clean.

In the next section (@sec-data-clean), we will outline the data-cleaning process and also show the structure of the cleaned data.

## Cleaned Data {#sec-data-clean}

As stated above in @sec-data-raw, we start with the first dataset on GHS indices for all countries in the world. @tbl-cleaned-ghs shows the first 6 countries and their GHS indices. Here we only have 2 variables, namely the Country Name and the respective GHS index. We cleaned this dataset by selecting only the two columns for the respective variables and filtering the data for the year of 2021.

```{r}
#| echo: false
#| eval: true
#| label: tbl-cleaned-ghs
#| tbl-cap: "Cleaned Data showing GHS indices for first 6 countries"
#| tbl-cap-location: bottom
#| warning: false

cleaned_ghs_index_data_2021 <- read_csv("../data/analysis_data/cleaned_GHS_index.csv", show_col_types = FALSE)

table_cleaned_ghs_index_data_2021 <- cleaned_ghs_index_data_2021 |>
  rename(
    `Country Name` = country,
    `GHS Index` = overall_score
  )

kable(head(table_cleaned_ghs_index_data_2021))
```

Second, we have a raw dataset on all cause death rates for all countries from 1960-2021. We cleaned this dataset by selecting a subset of countries that helped us analyse how Russia fared against other countries in death ratesover the years. @tbl-cleaned-death-rates shows the structure of the cleaned data after we filtered it for the time period of 2010-2021 for India, South Korea, Russia and the United States.

```{r}
#| echo: false
#| eval: true
#| label: tbl-cleaned-death-rates
#| tbl-cap: "Cleaned Data showing all cause death rates for 4 countries"
#| tbl-cap-location: bottom
#| warning: false

cleaned_deaths_data <- read_csv("../data/analysis_data/cleaned_all_cause_dth_rates.csv", show_col_types = FALSE)

cleaned_deaths_data$country_name[cleaned_deaths_data$country_name == "Russian Federation"] <- "Russia"
cleaned_deaths_data$country_name[cleaned_deaths_data$country_name == "Korea, Rep."] <- "South Korea"

table_cleaned_deaths_data <- cleaned_deaths_data |>
  rename(
    `Country Name` = country_name,
  )

kable(head(table_cleaned_deaths_data))
```

The final raw dataset, as mentioned in @sec-data-raw, that we have is the IHME data on the COVID death rate, excess deaths, excess death rate and ratio of excess death rate over COVID death rate for all countries in the world. In this paper we only needed the two variables excess death rate and the country name, however for better analysis of Russia's ability to control the pandemic's effect, we combined this data with the respective GHS index as shown in @tbl-cleaned-combined. Hence, there are three variables, the country name, and the respective excess death rate and GHS index.

```{r}
#| echo: false
#| eval: true
#| label: tbl-cleaned-combined
#| tbl-cap: "Cleaned Data showing Excess Death Rate and GHS index for first 6 chosen countries"
#| tbl-cap-location: bottom
#| warning: false

combined_data <- read_csv("../data/analysis_data/cleaned_excess_death_ghs.csv", show_col_types = FALSE)

table_combined_data <- combined_data |>
  rename(
    `Country Name` = country,
    `Excess Death Rate` = excess_death_rate,
    `GHS Index` = ghs_index
  )

kable(head(table_combined_data))
```

Now that we have cleaned our raw datasets and have generated clean tables for analysis, we will go on to the Results Section (@sec-results) and delve deeper into studying the effect of the COVID-19 pandemic on Russia and if it was impacted significantly more than other countries.

## Measurement {#sec-data-measurement}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aenean tempus libero sit amet nisl dapibus, varius dictum magna pellentesque. In lacinia nec ligula eget condimentum. Aenean imperdiet enim sit amet felis volutpat, eu ornare ante pretium. Donec ullamcorper tempus facilisis. Mauris porttitor libero ultrices tortor facilisis, quis varius sapien ultricies. Pellentesque et metus vel quam mollis dignissim ultricies a dui. Cras sodales porttitor pulvinar. Proin vel nunc sit amet sapien venenatis pellentesque nec at felis.

# Results {#sec-results}

This section will be divided into three parts where all together we aim to look at Russia's capabilty according to the GHS index to tackle pandemics (@sec-results-ghs), then compare Russia's trend in all cause death rates and excess death rates with those of other countries like India, South Korea and the United states (@sec-results-death-rates) and finally study the relationship between the excess death rate and GHS index for Russia along with a subset of countries in the world (@sec-results-death-rates).

## Global Distribution of Global Health Security Index (GHS) in 2021 {#sec-results-ghs}

As explained in @sec-data-raw, the GHS index is an indicator of how well prepared a country is to tackle epidemics or pandemics like the COVID-19 outbreak. The GHS index gives one an idea of how the sound health care system  as well as the rules and regulations of a certain country are in the context of facing pandemics. Generally, the higher the GHS index of a country, the more well-prepared and secure they are. @fig-map shows a representation of the world map with countries shaded on the basis of their GHS indices. This map is a replication of figure 1 in Nuzzo and Ledesma paper [@citeRepPaper] that is being replicated in this report.

```{r}
#| label: fig-map
#| fig-cap: Global Distribution of Global Health Security (GHS) Index Scores in 2021
#| echo: false
#| warning: false
#| message: false
#| fig-height: 6
#| fig-width: 8
cleaned_ghs_index_data_2021 <- read_csv("../data/analysis_data/cleaned_GHS_index.csv", show_col_types = FALSE)

world <- ne_countries(scale = "medium", returnclass = "sf")

world_scores <- merge(world, cleaned_ghs_index_data_2021, by.x = "name", by.y = "country", all.x = TRUE)

world_scores$score_factor <- cut(
  world_scores$overall_score, 
  breaks = c(16, 25, 30, 35, 40, 45, 50, 60, 70, 76), 
  labels = c("16 to 25", "25 to 30", "30 to 35", "35 to 40", "40 to 45", "45 to 50", "50 to 60", "60 to 70", "70 to 76"), 
  include.lowest = TRUE)

ggplot(data = world_scores) +
  geom_sf(aes(fill = score_factor), color = "white") +
  scale_fill_manual(values = c("#990000", "#CC3333", "#FF6666", "#FF9966", "#FFCC99", "#99CCFF", "#6699CC", "#336699", "#003399")) +
  labs(fill = "GHS Indices") +
  theme_minimal()
```

We can observe here that there are countries like the United States and Australia whose GHS indices are one of the highest, implying that they were deemed extremely well prepared for epidemics. Then we see countries like Russia and China who were, although moderately when compared to the aforementioned countries, but still well prepared to tackle pandemics. Finally, we see countries such as Venezuela and Yemen which were said to be the least secure of all.

@fig-map provides this report with a background on how secure Russia was in tackling epidemics and pandemics, which as seen in the map, it was in a good position to face such outbreaks where numerous countries were deemed less prepared than Russia. This will aid us in drawing meaningful analyses throughout the paper, specifically in @sec-results-death-rates, sec-results-death-rates and @sec-discussion.

## Trend in Death Rates for Russia, India, South Korea and United States {#sec-results-death-rates}

Now we move on to analyzing the trends in the all cause death rates per 1000 population over 2010-2021, which includes COVID deaths as well as general deaths which were not caused by the outbreak, and the trend in the excess death rates per 100 population in 2020 and 2021. The excess death rates in this section were calculated by finding a difference between the expected death rate and the actual death rate for that year.

@fig-all-cause-death-rate shows us the trend for the all cause death rates per 1000 population over 2010-2021 for India, South Korea, Russia and the United States using the data presented in @tbl-cleaned-death-rates. We observe here that from the years 2010-2019, there is a general gradually decreasing trend in the death rates in the case of Russia and India, and a gradually increasing trend in the death rates in the case of South Korea and the United States. The dashed line represents the line of best fit for the trend from 2010-2019.

```{r}
#| label: fig-all-cause-death-rate
#| fig-cap: All Cause Death Rates per 1000 Population for India, South Korea, Russia and United States (2010-2021)
#| echo: false
#| warning: false
#| message: false
#| fig-height: 4
#| fig-width: 8
cleaned_deaths_data <- read_csv("../data/analysis_data/cleaned_all_cause_dth_rates.csv", show_col_types = FALSE)

cleaned_deaths_data$country_name[cleaned_deaths_data$country_name == "Russian Federation"] <- "Russia"
cleaned_deaths_data$country_name[cleaned_deaths_data$country_name == "Korea, Rep."] <- "South Korea"

cleaned_deaths_data_long <- pivot_longer(cleaned_deaths_data, cols = -country_name, names_to = "Year", values_to = "Value")

cleaned_deaths_data_long[,-1] <- lapply(cleaned_deaths_data_long[,-1], as.numeric)

cleaned_deaths_data_fit <- cleaned_deaths_data_long |>
  filter(Year >= 2010 & Year <= 2019)

ggplot(cleaned_deaths_data_long, aes(x = Year, y = Value, group = country_name, color = country_name)) +
  geom_line(aes(color = country_name)) +
  geom_point(aes(color = country_name)) +
  geom_smooth(data = cleaned_deaths_data_fit, method = "lm", se = FALSE, aes(color = country_name), linetype = "dashed", size = 0.5) +
  geom_smooth(data = cleaned_deaths_data_long, method = "lm", se = FALSE, aes(color = country_name), linetype = "dotted", size = 1) +
  theme_minimal() +
  scale_x_continuous(breaks = 2010:2021) +
  labs(title = "All Cause Death Rates by Country (2010-2021)", x = "Year", y = "All Cause Death Rates") +
  theme(legend.title = element_blank())
```

We see here in @fig-all-cause-death-rate that there is a sudden increase in the death rate in 2020 and 2021, and the increase in Russia's death rate is much higher than that of the other countries. The essential point to note here is that in the case of India, which has a GHS index lower than that of Russia, the rising trend observed in 2020 and 2021 was still not as excessive as in the case of Russia. The dotted line represents the line of best fit for the trend from 2010-2021. As we go from comparing Russia to South Korea, we see that in the case of South Korea, both the dashed (2010-2019) and the dotted (2010-2021) are very close together implying that the country did not experience an unanticipated number of deaths during COVID. However, in the case of Russia, we see the two lines diverging implying that there were unforeseen deaths during that period.

This observation leads us to @fig-excess-death-rate, where we present the excess death rate for the four countries for 2020 and 2021. As mentioned above, the excess death rates were calculated by extrapolating the dashed line of best-fit from years 2010-2019 in @fig-all-cause-death-rate representing the trend in all cause death rates to 2020 and 2021. This way we were able to find the expected death rate for each of the four countries and subsequently we were able to plot the difference between the expected death rate and the actual death rate called the excess death rate for 2020 and 2021 for all four countries in @fig-excess-death-rate.

```{r}
#| label: fig-excess-death-rate
#| fig-cap: Excess Death Rates per 1000 Population for India, South Korea, Russia and United States, 2020 and 2021
#| echo: false
#| warning: false
#| message: false
#| fig-height: 4
#| fig-width: 8

# Model: India
model_india <- lm(Value ~ Year, data = cleaned_deaths_data_fit |> filter(country_name == "India"))
# Predicting 2020 and 2021
coefficients_ind <- coef(model_india)
ind_2020 <- (coefficients_ind["Year"] * 2020) + coefficients_ind["(Intercept)"]
ind_2021 <- (coefficients_ind["Year"] * 2021) + coefficients_ind["(Intercept)"]
# Ratio
ind <- cleaned_deaths_data_long |>
  filter(country_name == "India")
excess_ind_2020 <- ind$Value[11] - ind_2020
excess_ind_2021 <- ind$Value[12] - ind_2021

# Model: Russia
model_rus <- lm(Value ~ Year, data = cleaned_deaths_data_fit |> filter(country_name == "Russia"))
# Predicting 2020 and 2021
coefficients_rus <- coef(model_rus)
rus_2020 <- (coefficients_rus["Year"] * 2020) + coefficients_rus["(Intercept)"]
rus_2021 <- (coefficients_rus["Year"] * 2021) + coefficients_rus["(Intercept)"]
# Ratio
rus <- cleaned_deaths_data_long |>
  filter(country_name == "Russia")
excess_rus_2020 <- rus$Value[11] - rus_2020
excess_rus_2021 <- rus$Value[12] - rus_2021

# Model: United States
model_us <- lm(Value ~ Year, data = cleaned_deaths_data_fit |> filter(country_name == "United States"))
# Predicting 2020 and 2021
coefficients_us <- coef(model_us)
us_2020 <- (coefficients_us["Year"] * 2020) + coefficients_us["(Intercept)"]
us_2021 <- (coefficients_us["Year"] * 2021) + coefficients_us["(Intercept)"]
# Ratio
us <- cleaned_deaths_data_long |>
  filter(country_name == "United States")
excess_us_2020 <- us$Value[11] - us_2020
excess_us_2021 <- us$Value[12] - us_2021

# Model: South Korea
model_sk <- lm(Value ~ Year, data = cleaned_deaths_data_fit |> filter(country_name == "South Korea"))
# Predicting 2020 and 2021
coefficients_sk <- coef(model_sk)
sk_2020 <- (coefficients_sk["Year"] * 2020) + coefficients_sk["(Intercept)"]
sk_2021 <- (coefficients_sk["Year"] * 2021) + coefficients_sk["(Intercept)"]
# Ratio
sk <- cleaned_deaths_data_long |>
  filter(country_name == "South Korea")
excess_sk_2020 <- sk$Value[11] - sk_2020
excess_sk_2021 <- sk$Value[12] - sk_2021

excess_deaths <- tibble(
  Year = c(2020, 2020, 2020, 2020, 2021, 2021, 2021, 2021),
  country_name = rep(c("India", "Russia", "United States", "South Korea"), times = 2),
  Value = c(excess_ind_2020, excess_rus_2020, excess_us_2020, excess_sk_2020, excess_ind_2021, excess_rus_2021, excess_us_2021, excess_sk_2021) |> round(2)
)

ggplot(excess_deaths, aes(x = country_name, y = Value, fill = as.factor(Year))) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7) +
  coord_flip() +  # Flip coordinates to make the bar plot horizontal
  theme_minimal() +
  labs(title = "Excess Death Rate by Country and Year", x = "Country", y = "Excess Death Rate") +
  scale_fill_manual(values = c("2021" = "cyan4", "2020" = "cadetblue2"), name = "Year") +
  geom_text(aes(label = Value), position = position_dodge(width = 0.8), hjust = -0.01, color = "black")
```

We see in @fig-excess-death-rate the massive difference between the excess death rates for Russia and the other countries. Comparing Russia to South Korea we see that, as seen in @fig-all-cause-death-rate, South Korea was very faintly affected by the COVID-19 pandemic, however Russia saw a sudden rise in deaths from 2019 to 2021. This shows us that the observations here are analogous to the plot in @fig-all-cause-death-rate. When we compare Russia with India, which as aforementioned was deemed a less secure country in the case of epidemics, we see that the excess death rates for Russia in both 2020 and 2021 are much higher than those for India. This observation will further aid us in making meaningful discussions in @sec-discussion.

## Relationship between the Excess Death Rate and Global Health Security Index (GHS) {#sec-results-death-rates}

Finally, we will now explore the relationship between the excess death rate seen above and the GHS index for a subset of countries per 100,000 population. @fig-excess-death-rate-ghs shows us a subset of countries scattered around the plot according to their excess death rate and the GHS index shown in the Appendix @tbl-appendix. The green line shows the general decreasing trend which implies that as the GHS index increases, the excess death rate tends to decrease, i.e. the excess death rate for more "secure" or well-prepared countries is generally lower.

Russia, labelled in red, is seen to be significantly higher than the line of best fit, which implies that Russia's excess death rate was much worse than it should have been. We observe that countries like Brazil and Indonesia, which have approximately the same GHS index as Russia, are nearer to the line of best fit than Russia. India, although having a lower GHS index than Russia is seen to have a much lower than general excess death rate.

```{r}
#| label: fig-excess-death-rate-ghs
#| fig-cap: Relationship between the Excess Death Rate and the Global Health Security Index (GHS) per 100,000 Population for a subset of countries, 2020-2021
#| echo: false
#| warning: false
#| message: false
#| fig-height: 5
#| fig-width: 8

combined_data <- read_csv("../data/analysis_data/cleaned_excess_death_ghs.csv", show_col_types = FALSE)

ggplot(combined_data, aes(x = ghs_index, y = excess_death_rate, label = country)) +
  geom_point(aes(color = ifelse(country %in% c("Russia"), "Highlight", "Normal"))) +
  scale_color_manual(values = c("Highlight" = "red", "Normal" = "black")) +
  guides(color = FALSE) +
  geom_text(vjust = 1.5, hjust = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "#669900") +
  labs(title = "Excess Death Rate vs GHS Index for a subset of countries",
       x = "Global Health Security Index (GHS)",
       y = "Excess Death Rate") +
  theme_minimal()
```

The implications from this figure will be further studied in @sec-discussion.

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and Further Scope {#sec-disc-weaknesses}

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

```{r}
#| echo: false
#| eval: true
#| label: tbl-appendix
#| tbl-cap: "Cleaned Data showing Excess Death Rate and GHS index for a subset of countries"
#| tbl-cap-location: bottom
#| warning: false

combined_data <- read_csv("../data/analysis_data/cleaned_excess_death_ghs.csv", show_col_types = FALSE)

table_combined_data <- combined_data |>
  rename(
    `Country Name` = country,
    `Excess Death Rate` = excess_death_rate,
    `GHS Index` = ghs_index
  )

kable(table_combined_data)
```

<!-- In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows... -->

<!-- In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...  -->

<!-- ```{r} -->
<!-- #| eval: true -->
<!-- #| echo: false -->
<!-- #| message: false -->
<!-- #| warning: false -->
<!-- #| label: fig-ppcheckandposteriorvsprior -->
<!-- #| layout-ncol: 2 -->
<!-- #| fig-cap: "Examining how the model fits, and is affected by, the data" -->
<!-- #| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"] -->

<!-- pp_check(first_model) + -->
<!--   theme_classic() + -->
<!--   theme(legend.position = "bottom") -->

<!-- posterior_vs_prior(first_model) + -->
<!--   theme_minimal() + -->
<!--   scale_color_brewer(palette = "Set1") + -->
<!--   theme(legend.position = "bottom") + -->
<!--   coord_flip() -->
<!-- ``` -->

<!-- ## Diagnostics -->

<!-- @fig-stanareyouokay-1 is a trace plot. It shows... This suggests... -->

<!-- @fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests... -->

<!-- ```{r} -->
<!-- #| echo: false -->
<!-- #| eval: true -->
<!-- #| message: false -->
<!-- #| warning: false -->
<!-- #| label: fig-stanareyouokay -->
<!-- #| fig-cap: "Checking the convergence of the MCMC algorithm" -->
<!-- #| fig-subcap: ["Trace plot", "Rhat plot"] -->
<!-- #| layout-ncol: 2 -->

<!-- plot(first_model, "trace") -->

<!-- plot(first_model, "rhat") -->
<!-- ``` -->

<!-- # Data {#sec-data} -->

<!-- Some of our data is of penguins (@fig-bills), from @palmerpenguins. -->

<!-- ```{r} -->
<!-- #| label: fig-bills -->
<!-- #| fig-cap: Bills of penguins -->
<!-- #| echo: false -->

<!-- ggplot(penguins, aes(x = island, fill = species)) + -->
<!--   geom_bar(alpha = 0.8) + -->
<!--   scale_fill_manual(values = c("darkorange","purple","cyan4"), -->
<!--                     guide = "none") + -->
<!--   theme_minimal() + -->
<!--   facet_wrap(~species, ncol = 1) + -->
<!--   coord_flip() -->
<!-- ``` -->

<!-- Talk more about it. -->

<!-- And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.) -->

<!-- ```{r} -->
<!-- #| label: fig-planes -->
<!-- #| fig-cap: Relationship between wing length and width -->
<!-- #| echo: false -->
<!-- #| warning: false -->
<!-- #| message: false -->

<!-- analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv")) -->

<!-- analysis_data |>  -->
<!--   ggplot(aes(x = width, y = length)) + -->
<!--   geom_point(alpha = 0.8) + -->
<!--   theme_minimal() + -->
<!--   labs(x = "Wing width (mm)", -->
<!--        y = "Wing length (mm)") -->
<!-- ``` -->

<!-- Talk way more about it.  -->

<!-- # Results {#sec-results} -->

<!-- Our results are summarized in @tbl-modelresults. -->

<!-- ```{r} -->
<!-- #| echo: false -->
<!-- #| eval: true -->
<!-- #| warning: false -->
<!-- #| message: false -->

<!-- library(rstanarm) -->

<!-- first_model <- -->
<!--   readRDS(file = here::here("models/first_model.rds")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| echo: false -->
<!-- #| eval: true -->
<!-- #| label: tbl-modelresults -->
<!-- #| tbl-cap: "Explanatory models of flight time based on wing width and wing length" -->
<!-- #| warning: false -->

<!-- modelsummary::modelsummary( -->
<!--   list( -->
<!--     "First model" = first_model -->
<!--   ), -->
<!--   statistic = "mad", -->
<!--   fmt = 2 -->
<!-- ) -->
<!-- ``` -->

\newpage

# References